---
title: "Semantic Similarity Final"
author: "Jess Boyle"
date: "2023-11-06"
output: 
  html_document: 
    toc: yes
    number_sections: yes
---

This includes the statistical analysis for the semantic similarity analysis of the NCTE transcripts completed in goggle collab. 

#set up 

```{r}
rm(list=ls(all=TRUE)) #clear memory
library(tidyverse) #calling tideverse package
library(ggplot2) #calling ggplot package
library(dplyr) #calling dplyr package
library(corrplot)
library(PerformanceAnalytics) #for chart.Correlation
```

# Read in Similiarty Scores and CLASS Data
```{r}
#Read in similarity scores from google collab
sim_scores <- read_csv("ncte_teacher_similarityscores.csv") %>% 
  select(OBSID, alltext_similar_score, cw_similar_score)

# Calculate mean similarity score for all_text_similar_score
mean_alltext_similar_score <- sim_scores %>%
  group_by(OBSID) %>%
  summarize(
    count = n(),
    mean_alltext_similar_score = mean(alltext_similar_score)
  ) 

# Calculate mean similarity score for cw_similar_score
mean_cw_similar_score <- sim_scores %>%
  group_by(OBSID) %>%
  summarize(
    count = n(),
    mean_cw_similar_score = mean(cw_similar_score)
  )

# Join mean similarity scores 
mean_sim_scores <- left_join(mean_alltext_similar_score, mean_cw_similar_score, by = "OBSID") %>% 
  select (OBSID, mean_alltext_similar_score, mean_cw_similar_score)

#Read in CLASS measure scores
class_scores <- read_csv("class_data.csv") %>% 
  select(1,13,17:18)

#join dataframes by teacher ID
ncte_data <- left_join(mean_sim_scores, class_scores, by = "OBSID") %>% 
  filter(!is.nan(instructional_dialogue))

str(ncte_data)

```

## Examining Descriptives to Use for Describing Data Set
```{r}
#Read in CLASS measure scores
class_scores_id <- read_csv("class_data.csv")

teacher_id <- read_csv("ncte_obsid_ncteid.csv")

#join dataframes by teacher ID
class_scores_teacher_id <- left_join(class_scores_id, teacher_id, by = "OBSID") %>% 
  filter(!is.nan(instructional_dialogue))

# how many teachers included
unique_numbers <- unique(class_scores_teacher_id$NCTETID)
unique_numbers

# Calculate mean similarity score for cw_similar_score
mean_cw_similar_score <- class_scores_teacher_id %>%
  group_by(NCTETID) %>%
  summarize(
    count = n(),
    mean_cw_similar_score = mean(OBSID)
  )

# Calculate average number of observations per teacher
average_observations <- class_scores_teacher_id %>%
  group_by(NCTETID) %>%
  summarize(OBSID = n()) %>%
  summarize(average_observations = mean(OBSID))

```
# Visualize Mean Similarity Scores with Violin Plots
```{r}
#Violin Plot for All Words
ggplot(mean_sim_scores, aes(x = "", y = mean_alltext_similar_score)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(width = 0.1) +
  stat_summary(fun = mean, geom = "text", 
               aes(label = paste("Mean: ", round(..y.., 2)), x = 0.85),
               vjust = -1, size = 3.5, color = "blue") +
  ylab("Mean Similarity Scores") +
  ggtitle("Violin Plot for Mean Similarity Scores of All Words")

# Violin plot for Content Words
ggplot(mean_sim_scores, aes(x = "", y = mean_cw_similar_score)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(width = 0.1) +
  stat_summary(fun = mean, geom = "text", 
               aes(label = paste("Mean: ", round(..y.., 2)), x = 0.85),
               vjust = -1, size = 3.5, color = "blue") +
  ylab("Mean Similarity Scores") +
  ggtitle("Violin Plot for Mean Similarity Scores of Content Words")

psych::describe(mean_sim_scores)
```


# Visualize Mean Scores with Histograms
```{r}
#Histogram
ggplot(mean_sim_scores, aes(mean_alltext_similar_score)) +   
  geom_histogram(binwidth = 0.1, col = "blue", fill = "blue", alpha = .5) +  
  xlab("Raw Similarity Scores") + #label the axes
  ylab("Frequency") +
  ggtitle("Histogram for Mean Similarity Scores of All Words") #title

#Histogram
ggplot(mean_sim_scores, aes(mean_cw_similar_score)) +   
  geom_histogram(binwidth = 0.1, col = "blue", fill = "blue", alpha = .5) +  
  xlab("Raw Similarity Scores") + #label the axes
  ylab("Frequency") +
  ggtitle("Histogram for Mean Similarity Scores of Content Words") #title

```


# Examine Relationship between Similarity Scores & CLASS Variables
```{r}
cor(ncte_data[2:6], use = "pairwise.complete.obs")

library(PerformanceAnalytics) #for chart.Correlation
chart.Correlation(ncte_data[2:6], histogram = TRUE, method = "pearson")
```


# Examine Raw Similarity Score Data with Histograms
```{r}
#Histogram
ggplot(sim_scores, aes(alltext_similar_score)) +   
  geom_histogram(binwidth = 0.1, col = "blue", fill = "blue", alpha = .5) +  
  xlab("Raw Similarity Scores of All Text") + #label the axes
  ylab("Frequency") +
  ggtitle("Histogram for Raw Similarity Scores of All Text") #title

#Histogram
ggplot(sim_scores, aes(cw_similar_score)) +   
  geom_histogram(binwidth = 0.1, col = "blue", fill = "blue", alpha = .5) +  
  xlab("Raw Similarity Scores") + #label the axes
  ylab("Frequency") +
  ggtitle("Histogram for Raw Similarity Scores of Content Words") #title

```

# Examine Raw Similarity Score Data with Violin Plots
```{r}
# Violin plot for Content Words
ggplot(sim_scores, aes(x = "", y = cw_similar_score)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(width = 0.1) +
  stat_summary(fun = mean, geom = "text", 
               aes(label = paste("Mean: ", round(..y.., 2)), x = 0.85),
               vjust = -1, size = 3.5, color = "blue") +
  ylab("Raw Similarity Scores") +
  ggtitle("Violin Plot for Raw Similarity Scores of Content Words")
```




# Remove Zero Scores in Content Words and Examine New Correlation with CLASS Outcomes
```{r}
#filter out 0 rows
filtered_scores <- sim_scores %>%
  filter(cw_similar_score != 0)
#Went from 207427 observations to 197009 (removed 10,000)

# Calculate mean similarity score for filtered cw_similar_score
filtered_mean_cw_sim_scores <- filtered_scores %>%
  group_by(OBSID) %>%
  summarize(
    count = n(),
    mean_cw_similar_score = mean(cw_similar_score)
  )

# Calculate mean similarity score for all_text_similar_score
filtered_mean_alltext_sim_scores <- filtered_scores %>%
  group_by(OBSID) %>%
  summarize(
    count = n(),
    mean_alltext_similar_score = mean(alltext_similar_score)
  ) 

# Join filtered mean similarity scores 
filtered_mean_sim_scores <- left_join(filtered_mean_alltext_sim_scores, filtered_mean_cw_sim_scores, by = "OBSID") %>% 
  select (OBSID, mean_alltext_similar_score, mean_cw_similar_score)

#join dataframes by teacher ID
filtered_ncte_data <- left_join(filtered_mean_sim_scores, class_scores, by = "OBSID") %>% 
  filter(!is.nan(instructional_dialogue))

```


# How are Filtered and Unfiltered Scores Different
```{r}
library(psych) #load in psych package

psych::describe(ncte_data)
psych::describe(filtered_ncte_data)

```

# Examine Relationship between Filtered Similarity Scores & CLASS Variables
```{r}
cor(filtered_ncte_data[2:6], use = "pairwise.complete.obs")

library(PerformanceAnalytics) #for chart.Correlation
chart.Correlation(filtered_ncte_data[2:6], histogram = TRUE, method = "pearson")

```







